A comparative analysis of penalized linear mixed models in structured genetic data

### 2021-02-15 main.tex
Sec 1 - Responding to reviewer comment "More  background  on  lasso"

One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
%LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously. However, classical multivariate approaches rely on data with more observations than unknown parameters. This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals. Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates. The lasso is a widely used penalized regression method \citep{tibshirani1996regression}. \anna{More background on lasso} It relies on an assumption of sparsity; in other words, that many features will have no effect. Thus, although the total number of features may be large, the number that actually affect the outcome will be small. This allows coefficient estimation and variable selection to be accomplished simultaneously, and provides considerable computational advantages that scale up well to very large data sets. These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest. Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.

@@

All analyses and simulations were performed using \texttt{R} version ???. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix was implemented using \texttt{ggmix} \citep{bhatnagar2020simultaneous}. The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}. Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$. \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.

>>>>

One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.

An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously. However, classical multivariate approaches rely on data with more observations than unknown parameters. This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals.

Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates. \anna{\textbf{More background on lasso}} The lasso is a widely used penalized regression method \citep{tibshirani1996regression}, \anna{which can be used to model continuous, dichotomous, and survival outcomes.}  It relies on an assumption of sparsity; in other words, that many features will have no effect \anna{so that their coefficients will be estimated to be exactly 0}. Thus, although the total number of features may be large, the number that actually affect the outcome will be small. This allows coefficient estimation and variable selection to be accomplished simultaneously, and provides considerable computational advantages that scale up well to very large data sets. \anna{For the small number of features that have an effect on the outcome, large, extreme coefficient magnitudes are assumed to be less realistic and are thus shrunken towards 0.} These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest, \anna{and the effect size of a variant is typically assumed small}. Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.

@@

All analyses and simulations were performed using \texttt{R} version 3.6.3. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix was implemented using \texttt{ggmix} \citep{bhatnagar2020simultaneous}. The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}. Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$. \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
